Sampling requirements to detect ecosystem change

### 2019-11-21 R/temporal-sampling.Rmd
added theoretical results

par(mfrow=c(2,2))
# Sigma plot
params <- tibble(sigma=rep(seq(0.1,20,0.1),100))

@@

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(sigma))) %>%
  plot(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1)

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(phi) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(phi))) %>%
  plot(ylab='Probability correct',xlab='phi',pch=16,cex.lab=1.3,las=1)

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_size) %>%
  summarize(prob_correct = sum(detection==1)/length(unique(shift_size))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1)

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = 20, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_time) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_time/unique(shift_time))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1)

>>>>

par(mfrow=c(2,2))
# Sigma plot
sample_size=10
params <- tibble(sigma=rep(seq(0.1,20,0.1),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 5, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(sigma/unique(sigma))) %>%
  plot(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

sample_size=10
params <- tibble(sigma=rep(seq(0.1,20,0.1),10))

@@

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(sigma/unique(sigma))) %>%
  points(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col='red')

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(phi) %>%
  summarize(prob_correct = sum(detection==1)/length(phi/unique(phi))) %>%
  plot(ylab='Probability correct',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_size) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_size/unique(shift_size))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

@@

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_time) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_time/unique(shift_time))) %>%
  plot(ylab='Probability correct',xlab='Shift timing',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

In line with theory about monitoring, we found that the probability of correcting identifying a changepoint decreased with larger levels of population variability ($\\sigma$) and autocorrelation ($\\phi$). We also found that the probability of correct changepoint detection increases with larger shift sizes.

The probability of correcting identifying a changepoint, of course, depends strongly on the number of samples.

```{r different_methods, fig.cap='Theoretical results here',message=F,warning=F,echo=F}
# 4(or 6?) panel plot

par(mfrow=c(1,1))
# Sigma plot
sample_size=10
params <- tibble(sigma=rep(seq(0.1,20,0.1),100))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 5, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(sigma/unique(sigma))) %>%
  plot(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

sample_size=10
params <- tibble(sigma=rep(seq(0.1,20,0.1),10))

output %>%
  group_by(sigma) %>%
  summarize(prob_correct = sum(detection==1)/length(sigma/unique(sigma))) %>%
  points(ylab='Probability correct',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col='red')

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(phi) %>%
  summarize(prob_correct = sum(detection==1)/length(phi/unique(phi))) %>%
  plot(ylab='Probability correct',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_size) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_size/unique(shift_size))) %>%
  plot(ylab='Probability correct',xlab='shift size',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))

output <- params %>%
  pmap(build_time_series) %>%
  map_dbl(function(df) ut(y=df, sample_size = sample_size, threshold_error = 3, each = 1, type="mean",random_sampling = F,convergence_sampling = F,regular_sampling = T)$regular_cptmean$same_cpt) %>%
  mutate(params,detection=.)

output %>%
  group_by(shift_time) %>%
  summarize(prob_correct = sum(detection==1)/length(shift_time/unique(shift_time))) %>%
  plot(ylab='Probability correct',xlab='Shift timing',pch=16,cex.lab=1.3,las=1,ylim=c(0,1))
